Equalizer\\Programming and User Guide

### 2012-02-09 Developer/ProgrammingGuide/paper.tex
Finalize Eq 1.2 PUG

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications using
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.8.6 for Equalizer 1.2\\[\medskipamount]
%  \today
}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.8     & Mar 21, 2011 & Revision for Equalizer 1.0\\
    1.8.1   & Mar 25, 2011 & Minor Revision\\
    1.8.2   & Apr 04, 2011 & Added \sref{sFocusDistance}\\
    1.8.3   & Aug 10, 2011 & Rewrote \sref{sHello}\\
    1.8.4   & Nov 18, 2011 & Added \sref{sAutoConfig}\\
    1.8.5   & Jan 27, 2012 & RC review for Equalizer 1.2\\
    1.8.6   & Feb 01, 2012 & RC2, added to \sref{sThreadModel}\\
    \hline
  \end{tabularx}}
\clearpage

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2em}
  \includegraphics[width=.618\textwidth]{images/DPlex.pdf}
  {\caption{ A DPlex Compound}}
\end{wrapfigure}
DPlex compounds assign full, alternating frames to individual rendering
units. The resulting images are copied to the destination channel, and
Equalizer load-balancing is used to ensure a steady framerate on the
destination window. This mode is also known as time-multiplex or AFR.

@@

@@

Pixel compounds are similar to 2D compounds. While 2D compounds decompose the
screen space into contiguous regions, pixel compounds assign one pixel of a
regular kernel to each resource. The frusta of the source rendering units are
modified so that each unit renders an evenly distributed subset of pixels, as
shown in \fig{fPixel}\footnote{3D model courtesy of AVS, USA.}

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/Pixel.pdf}
  {\caption{\label{fPixel}\small A Pixel Compound}}
\end{wrapfigure}
As 2D compounds, pixel compounds have low, constant IO requirements for the
pixel transfers during recomposition. They are naturally load-balanced for
fill-limited operations, but do not scale geometry processing at all.

@@

Subpixel compounds are configured using the \textsf{subpixel} parameter, using
the values \textsf{[~index size ]}. The \textsf{index} parameter corresponds to
the current resource index and the \textsf{size} is the total number of
resources used.

@@

\subsubsection{GPU Service Discovery}
Equalizer uses the GPU-SD library for discovering local and remote GPUs. Based
on this information, a typical configuration is dynamically compiled at
application start. The auto-configuration uses the same format as the static
configuration files, and can therefore be used to create template configurations
by running a test application and copying the generated configuration.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/gpu-sd.pdf}
  {\caption{\label{fGPUSD}GPU discovery for auto-configuration}}
\end{wrapfigure}
GPU-SD contains three components: a core library, local and remote discovery
modules as well as a ZeroConf daemon. The daemon uses the appropriate module to
query local GPUs and announces them using ZeroConf on the local network.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frusta.pdf}
  {\caption{\label{fFrusta}Wall and Projection Parameters}}
\end{wrapfigure}
\fig{fFrusta} illustrates the wall and projection frustum parameters.

@@

A canvas has one or more layouts. One of the layouts is the active
layout, that is, this set of views is currently used for rendering. It
is possible to specify \textsf{OFF} as a layout, which deactivates the
canvas. It is possible to use the same layout on different canvases.

>>>>

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications using
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.10 for Equalizer 1.2\\[\medskipamount]
%  \today
}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.8     & Mar 21, 2011 & Revision for Equalizer 1.0\\
    1.10    & Feb 09, 2012 & Revision for Equalizer 1.2\\
    \hline
  \end{tabularx}}
\clearpage

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}
  \includegraphics[width=.618\textwidth]{images/DPlex.pdf}
  {\caption{ A DPlex Compound}}
\end{wrapfigure}
DPlex compounds assign full, alternating frames to individual rendering
units. The resulting images are copied to the destination channel, and
Equalizer load-balancing is used to ensure a steady framerate on the
destination window. This mode is also known as time-multiplex or AFR.

@@

\subsection{\label{sTileCompunds}Tile Compounds}

Tile compounds are similar to 2D compounds. They decompose the rendering in
screen-space, where each rendering unit pulls and processes regular tiles of the
final view. Tile compounds are ideal for purely fill-limited applications such
as volume rendering and raytracing.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/tile.pdf}
  {\caption{\label{fPixel}\small A Tile Compound}}
\end{wrapfigure}
Tile compounds have a low, constant IO overhead for the image transfers and can
provide good scalability when used with fill-limited applications. Tile
decomposition works transparently for all Equalizer-based applications.

Contrary to all other compounds, the work distribution is not decided
before-hand using a push model, but each resource pulls work from a central work
queue, until all tiles have been processed. Therefore the rendering is naturally
load-balanced since all sources pull data on an as-needed basis. The queue
implementation uses prefetching to hide the communication overhead of the
central, distributed tile queue.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}\includegraphics[width=.618\textwidth]{images/Pixel.pdf}
  {\caption{\label{fPixel}\small A Pixel Compound}}\vspace{-8ex}
\end{wrapfigure}
Pixel compounds are similar to 2D compounds. While 2D compounds decompose the
screen space into contiguous regions, pixel compounds assign one pixel of a
regular kernel to each resource. The frusta of the source rendering units are
modified so that each unit renders an evenly distributed subset of pixels, as
shown in \fig{fPixel}\footnote{3D model courtesy of AVS, USA.}

As 2D compounds, pixel compounds have low, constant IO requirements for the
pixel transfers during recomposition. They are naturally load-balanced for
fill-limited operations, but do not scale geometry processing at all.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/Subpixel.pdf}
  {\caption{\label{fSubpixel}\small A Subpixel Compound}}
\end{wrapfigure}
Subpixel compounds are configured using the \textsf{subpixel} parameter, using
the values \textsf{[~index size ]}. The \textsf{index} parameter corresponds to
the current resource index and the \textsf{size} is the total number of
resources used.

@@

\subsubsection{GPU Service Discovery}

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/gpu-sd.pdf}
  {\caption{\label{fGPUSD}GPU discovery for auto-configuration}}
\end{wrapfigure}
Equalizer uses the GPU-SD library for discovering local and remote GPUs. Based
on this information, a typical configuration is dynamically compiled at
application start. The auto-configuration uses the same format as the static
configuration files, and can therefore be used to create template configurations
by running a test application and copying the generated configuration.

@@

GPU-SD contains three components: a core library, local and remote discovery
modules as well as a ZeroConf daemon. The daemon uses the appropriate module to
query local GPUs and announces them using ZeroConf on the local network.

@@

\fig{fFrusta} illustrates the wall and projection frustum parameters.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frusta.pdf}
  {\caption{\label{fFrusta}Wall and Projection Parameters}}
\end{wrapfigure}
A canvas has one or more layouts. One of the layouts is the active
layout, that is, this set of views is currently used for rendering. It
is possible to specify \textsf{OFF} as a layout, which deactivates the
canvas. It is possible to use the same layout on different canvases.
